# Exploration {.unnumbered}

```{r}
#| label: setup
#| message: false
#| warning: false

library(here)
library(tidyverse)
library(lubridate)
library(scales)
library(units)
library(janitor)
library(glue)
library(readxl)
library(gt)
library(patchwork)
library(broman)
library(countrycode)

theme_set(theme_light() + 
            theme(panel.border = element_blank(),
                  panel.grid = element_blank(),
                  plot.title = element_text(size = rel(2.0),
                                            face = "bold"),
                  plot.title.position = "plot"
            )
)

my_caption <- "Daniel Moul. Source: US Peace Corps\nvia Data Liberation Project FOI request"

n_cutoff <- 50

```

```{r}

###### Prepare some extra categorical information

df_reasons_primary <- bind_rows(
  tibble(
    descr_primary = c("personal/family", "other personal/family related", "friend(s) or other family responsibility", 
                      "death/illness of close friend or family member", "other family", "family financial obligations"),
    reason_primary = "personal/family"
  ),
  tibble(
    descr_primary = c("adaptation", "unrealistic expectations", "emotional/mental health", "host country culture", 
                      "preferred another country/region", "homesick"),
    reason_primary = "adaptation/mental health"
  ),
  tibble(
    descr_primary =  c("resignation in lieu of administrative separation", "peace corps policies", "illegal substance", 
                       "alcohol", "unable to adhere to peace corps policies", "use of illegal substance or misuse of alcohol"),
    reason_primary = "policy"
  ),
  tibble(
    descr_primary = c("other career opportunity", "further education", "financial"),
    reason_primary = "career/finances"
  ),
  tibble(
    descr_primary = c("job related", "matching skills with work assignment", "other program/work assignment", 
                      "other program/work related (please explain)", "host country counterpart/support", 
                      "other country assignment/adaptation", "amount of work", 
                      "project/site development", "language skills", "host community/host family", "site location", 
                      "technical skills", "technical support", "living arrangements/housing location", 
                      "assignment not development work/not needed", "insufficient work to do",
                      "other job related issues (philosophical difference)", "received inaccurate information re: assignment"),
    reason_primary = "assignment"
  ),
  tibble(
    descr_primary = c("physical health", "medical concerns", "medical support") ,
    reason_primary = "health"
  ),
  tibble(
    descr_primary = c("personal safety", "other personal safety", "crime and personal safety", "sexual assault/harassment", 
                      "political/civil unrest", "road safety/traveling", "other personal safety (please explain)", 
                      "physical assault/harassment", "political/civil unrest makes work too difficult", 
                      "road safety concerns/dangerous traveling", "other concerns related to personalsafety"),
    reason_primary = "safety"
  ),
  tibble(
    descr_primary = c("other", "prefer not to respond", "other (please explain)", "v/t prefers not to respond"),
    reason_primary = "other"
  ),
  tibble(
    descr_primary = c("romantic interest", "spousal responsibility", "early termination of spouse", 
                      "divorce or separation from pcv/t spouse"),
    reason_primary = "partner"
  ),
  tibble(
    descr_primary = c("other volunteers/trainees", "administrative support", "other peace corps support",
                      "peace corps staff in country", "conflict with volunteers/trainees"),
    reason_primary = "peace corps"
  )
) |>
  select(reason_primary, descr_primary) |>
  arrange(reason_primary, descr_primary)

df_reasons_secondary <- bind_rows(
  tibble(
    descr_secondary = c("death/illness of close friend or family member", "family financial obligations", 
      "friend(s) or other family responsibility", "other family", "other personal/family related", 
      "personal/family"),
    reason_secondary = "personal/family"
  ),
  tibble(
    descr_secondary = c("adaptation", "emotional/mental health", "host country culture", "language skills", 
      "other country assignment/adaptation", "preferred another country/region", 
      "unrealistic expectations"),
    reason_secondary = "adaptation/mental health"
  ),
  tibble(
    descr_secondary = c("alcohol", "illegal substance", "peace corps policies", "resignation in lieu of administrative separation", 
      "unable to adhere to peace corps policies", "use of illegal substance or misuse of alcohol"),
    reason_secondary = "policy"
  ),
  tibble(
    descr_secondary = c("financial", "further education", "other career opportunity"),
    reason_secondary = "career/finances"
  ),
  tibble(
    descr_secondary = c("amount of work", "assignment not development work/not needed", "conflicts with hc supervisor", 
      "host community/host family", "host country counterpart/support", "job not related to future career goals", 
      "job related", "living arrangements/housing location", "matching skills with work assignment", 
      "other job related issues (philosophical difference)", "other program/work assignment", 
      "other program/work related (please explain)", "project/site development", "received inaccurate information re: assignment", 
      "site location", "technical skills", "technical support") ,
    reason_secondary = "assignment"
  ),
  tibble(
    descr_secondary = c("medical concerns", "medical support"),
    reason_secondary = "health"
  ),
  tibble(
    descr_secondary = c("crime and personal safety", "other personal safety", "other personal safety (please explain)", "personal safety", 
      "physical assault/harassment", "physical health", "political/civil unrest", "political/civil unrest makes work too difficult", 
      "road safety concerns/dangerous traveling", "road safety/traveling", "sexual assault/harassment"),
    reason_secondary = "safety"
  ),
  tibble(
    descr_secondary = c("other", "other (please explain)", "prefer not to respond", "v/t prefers not to respond"),
    reason_secondary = "other"
  ),
  tibble(
    descr_secondary = c("divorce or separation from pcv/t spouse", "early termination of spouse", 
      "romantic interest", "spousal responsibility"),
    reason_secondary = "partner"
  ),
  tibble(
    descr_secondary = c("administrative support", "conflict with volunteers/trainees", "other peace corps support", 
      "other volunteers/trainees", "peace corps staff in country"),
    reason_secondary = "peace corps"
  )
) |>
  select(reason_secondary, descr_secondary) |>
  arrange(reason_secondary, descr_secondary)

df_country_info <- countrycode::codelist |>
  clean_names() |>
  select(country_name_en, continent, region) |>
  mutate(country_name_en = str_to_lower(country_name_en))


###### Get the data

dta <- read_xlsx(here("data/raw/Attachment_D_data_pulled_7_31_23.xlsx")) |>
  clean_names()  |>
  mutate(across(where(is.character), str_to_lower)) |>
  rename(descr_primary = et_prim_desc,
         descr_secondary = et_sec_desc,
         count = count_of_vol_id) |>
  separate_wider_delim(post_name, names = c("post", "city"),
                       delim = "/",
                       too_few = "align_start") |>
  select(-c(et_desc, city, et_prim_reason, et_sec_reason)) |>
  left_join(df_reasons_primary,
             by = "descr_primary") |>
  left_join(df_reasons_secondary,
             by = "descr_secondary") |>
  left_join(df_country_info,
            by = join_by(post == country_name_en)) |>
  mutate(region = if_else(post %in% c("png", "micronesia", "myanmar"),
                          "East Asia & Pacific",
         region),
         region = if_else(post %in% c("car", "congo", "cote d ivoire", "zaire", "sao tome", "guinea bissau"),
                          "Sub-Saharan Africa",
         region),
         region = if_else(post %in% c("czech", "russia far east", "russia western", "bosnia"),
                          "Europe & Central Asia",
         region),
         region = if_else(post %in% c("eastern caribbean"),
                          "Latin America & Caribbean",
         region)
  ) |>
  filter(!region == "pcr us relief") # leaving out 19 data points; since I don't know what this is (which region?)

```

```{r}

###### prepare dataframes

d_totals_post <- dta |>
  filter(is.na(descr_primary)) |>
  remove_empty(which = "cols")

top_post_subset <- d_totals_post |>
  slice_max(order_by = count, n = n_cutoff)

d_totals_post_primary <- dta |>
  filter(!is.na(descr_primary) & is.na(descr_secondary)) |>
  remove_empty(which = "cols") |>
  count(post, region, reason_primary, wt = count, sort = TRUE) |>
  mutate(pct_in_post = n / sum(n),
         .by = post)

d_totals_region_primary <- dta |>
  filter(!is.na(descr_primary) & is.na(descr_secondary)) |>
  remove_empty(which = "cols") |>
  count(region, reason_primary, wt = count, sort = TRUE) |>
  mutate(pct_in_region = n / sum(n),
         .by = region)

d_totals_primary <- dta |>
  filter(!is.na(descr_primary) & is.na(descr_secondary)) |>
  remove_empty(which = "cols") ##|>
  ##count(reason_primary, wt = count, sort = TRUE)

d_totals_primary_reason <- d_totals_primary |>
  count(reason_primary, wt = count, sort = TRUE)

d_totals_secondary_reason <- dta |>
  filter(!is.na(descr_secondary)) |>
  count(reason_secondary, wt = count, sort = TRUE)

```

## Posts by region

Since some small missions cover multiple countries, Peace Corps uses the term "post" instead of "country". There are `r nrow(d_totals_post)` posts in the data set.

Below are the `r n_cutoff` posts with the most resignations. See @fig-all-posts-count for the full list.

::: {.callout-warning}
## Be careful with comparisons

Without knowing how many volunteers went to each country in this time period, simple counts don't help us to assess which countries are under- or over-represented in resignations.

A more valid comparison is the *proportion of reasons* and *proportion of regions* in the plots with "B" panels below.

:::

```{r fig.height=6, fig.width=8}
#| label: fig-most-resignations-post-region
#| fig-cap: "Posts with most Peace Corps resignations, grouped by region"
#| fig-height: 6
#| fig-width: 8
#| column: page-right

d_totals_post |>
  mutate(post = fct_lump_n(post, n_cutoff, w = count)) |>
  count(post, region, wt = count, sort = TRUE) |>
  mutate(pct_of_all = n / sum(n)) |>
  filter(post != "Other") |>
  mutate(post = fct_reorder(post, n)) |>
  ggplot(aes(n, post, fill = region)) +
  geom_col(show.legend = FALSE,
           alpha = 0.7) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +
  scale_fill_viridis_d(end = 0.9) +
  facet_wrap(~region, scales = "free_y") +
  labs(
    title = glue("Posts with most Peace Corps volunteer resignations"),
    subtitle = glue("Top {n_cutoff} posts by region"),
    x = NULL,
    y = NULL,
    caption = my_caption
  )

```

<br>

## Consolidated reasons

The spreadsheet provided by the Peace Corps includes too many categories. So I consolidated them into 10 categories (@fig-reasons). The detailed mapping is available in @tbl-reasons-mapping. 

```{r fig.height=6, fig.width=12}
#| label: fig-reasons
#| fig-cap: "Primary reasons"
#| fig-height: 5
#| fig-width: 7
#| column: page-right

dta_for_plot <- d_totals_post_primary |>
  count(reason_primary, wt = n, sort = TRUE) |>
  mutate(pct_of_all = n / sum(n)) |>
  mutate(reason_primary = fct_reorder(reason_primary, n, sum)) |>
  arrange(desc(reason_primary)) |>
  mutate(end_pct = cumsum(pct_of_all),
         start_pct = lag(end_pct, default = 0)
         )

dta_for_plot |>
  ggplot() +
  geom_errorbarh(aes(xmin = start_pct, xmax = end_pct, y = reason_primary, color = reason_primary),
                 height = 0.3,
               show.legend = FALSE,
               alpha = 0.7) +
  scale_x_continuous(expand = expansion(mult = c(0.005, 0.05)),
                     labels = label_percent(),
                     breaks = 0:5 * 0.2) +
  scale_color_viridis_d(end = 0.9) +
  theme(axis.ticks = element_blank()) +
  labs(
    title = glue("Proportions of reasons"),
    subtitle = glue("All posts"),
    x = NULL,
    y = NULL,
    caption = my_caption
  )

```

<br>

## Regions and reasons

```{r fig.height=6, fig.width=10}
#| label: fig-most-resignations-reasons-by-region
#| fig-cap: "Count and proportion of resignations reasons by region"
#| fig-height: 6
#| fig-width: 10
#| column: page-right

p1 <- d_totals_region_primary |>
  mutate(region = fct_reorder(region, n),
         reason_primary = fct_reorder(reason_primary, pct_in_region, sum)) |>
  ggplot(aes(n, region, fill = reason_primary)) +
  geom_col(show.legend = TRUE,
           alpha = 0.7) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +
  # scale_fill_viridis_d(end = 0.9) +
  theme(legend.position = c(0.7, 0.3)) +
  labs(
    subtitle = "A: Count",
    x = NULL,
    y = NULL
  )

p2 <- d_totals_region_primary |>
  mutate(region = fct_reorder(region, n),
         reason_primary = fct_reorder(reason_primary, pct_in_region, sum)) |>
  ggplot(aes(pct_in_region, region, fill = reason_primary)) +
  geom_col(show.legend = FALSE,
           alpha = 0.7) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = label_percent()) +
  # scale_fill_viridis_d(end = 0.9) +
  labs(
    subtitle = "B: Relative proportion in region",
    x = NULL,
    y = NULL
  )

p1 + p2 +
  plot_annotation(
    title = glue("Regions by reason"),
    subtitle = glue("Count and relative proportion of reasons; includes all posts"),
    caption = my_caption
  )

```

<br>

A note re: @fig-most-resignations-reason-by-region: since there are so few data points from South Asia compared to the other regions (@fig-most-resignations-reasons-by-region), the large proportions of "other" and "health" reasons in South Asia below may not be as noteworthy as they first appear.

```{r fig.height=6, fig.width=10}
#| label: fig-most-resignations-reason-by-region
#| fig-cap: "Count and proportion of reasons colored by region"
#| fig-height: 6
#| fig-width: 10
#| column: page-right

p1 <- d_totals_region_primary |>
  mutate(region = fct_reorder(region, n),
         reason_primary = fct_reorder(reason_primary, pct_in_region, sum)) |>
  ggplot(aes(pct_in_region, reason_primary, fill = region)) +
  geom_col(show.legend = TRUE,
           alpha = 0.7) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +
  scale_fill_viridis_d(end = 0.9) +
  theme(legend.position = c(0.7, 0.3)) +
  labs(
    subtitle = "A: Count",
    x = NULL,
    y = NULL
  )

p2 <- d_totals_region_primary |>
  mutate(region = fct_reorder(region, n),
         reason_primary = fct_reorder(reason_primary, pct_in_region, sum)) |>
  ggplot(aes(pct_in_region, reason_primary, fill = region)) +
  geom_col(show.legend = FALSE,
           position = position_fill(),
           alpha = 0.7) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = label_percent()) +
  scale_fill_viridis_d(end = 0.9) +
  labs(
    subtitle = "B: Relative proportion in region",
    x = NULL,
    y = NULL
  )

 p1 + p2 +
  plot_annotation(
    title = glue("Reasons by region"),
    subtitle = glue("Count and relative proportion of regions for each reason; includes all posts"),
    caption = my_caption
  )

```

<br>

## Reasons ranked

The data set includes primary and secondary reasons in most cases. I use the same categorization for both. As noted in @fig-primary-secondary-reasons-count, number of resignations differ (1) in total between primary and secondary reasons; and (2) in the counts of the categories.

```{r fig.height=8, fig.width=10}
#| label: fig-primary-secondary-reasons-count
#| fig-cap: "Ranking and counts of primary and secondary reasons for resignation"
#| fig-height: 8
#| fig-width: 10
#| column: page-right
#| warning: false

dta_for_plot <- bind_rows(
  d_totals_primary_reason |>
    mutate(source = "primary") |>
    rename(reason = reason_primary),
  d_totals_secondary_reason  |>
    mutate(source = "secondary") |>
    rename(reason = reason_secondary)
) %>%
  mutate(rank = rank(n),
         source_label = glue("{cur_group()} (n={sum(n)})"),
         .by = source) |>
  mutate(reason = str_replace_all(reason, "adaptation/mental health", "adaptation/\nmental health"))

p1 <- dta_for_plot |>
  ggplot(aes(source_label, rank , color = reason, group = reason)) +
  geom_point(size = 3, alpha = 1) +
  geom_line() +
  geom_label(aes(label = if_else(str_detect(source_label, "primary"),
                                 reason,
                                 NA_character_)),
                 hjust = 1,
                 na.rm = TRUE) +
  geom_label(aes(label = if_else(str_detect(source_label, "secondary"),
                                 reason,
                                 NA_character_)),
                 hjust = 0,
                 na.rm = TRUE) +
  guides(color = "none") +
  scale_x_discrete(expand = expansion(mult = c(0.75, 0.75)),
                   position = "top") +
  scale_y_continuous(breaks = 0:5 * 2) +
  labs(
    subtitle = "A: Ranked (1 is lowest)",
    x = NULL,
    y = "Rank"
  )

p2 <- dta_for_plot |>
  ggplot(aes(source_label, n, color = reason, group = reason)) +
  geom_point(size = 3, alpha = 1) +
  geom_line() +
  geom_label(aes(label = if_else(str_detect(source_label, "primary"),
                                 reason,
                                 NA_character_)),
                 hjust = 1,
                 na.rm = TRUE) +
  geom_label(aes(label = if_else(str_detect(source_label, "secondary"),
                                 reason,
                                 NA_character_)),
                 hjust = 0,
                 na.rm = TRUE) +
  guides(color = "none") +
  scale_x_discrete(expand = expansion(mult = c(0.75, 0.75)),
                   position = "top") +
  expand_limits(x = 0) +
  labs(
    subtitle = "B: Count",
    x = NULL,
    y = "Count",
  )

p1 + p2 +
  plot_annotation(
    title = "Primary and secondary reasons\nfor Peace Corps volunteers' resignations",
    caption = my_caption
  )

```

<br>

```{r}
#| label: tbl-reasons
#| tbl-cap: "Reasons"
#| layout-ncol: 2

d_totals_primary_reason |>
  mutate(reason_primary = str_replace_all(reason_primary, "adaptation/mental health", "adaptation/\nmental health")) |>
  adorn_totals(where = "row") |>
  gt() |>
  tab_header(md("**Count of primary reasons**"))

d_totals_secondary_reason |>
  mutate(reason_secondary = str_replace_all(reason_secondary, "adaptation/mental health", "adaptation/\nmental health")) |>
  adorn_totals(where = "row") |>
  gt() |>
  tab_header(md("**Count of secondary reasons**"))

```

<br>

## Appendix

### Category consolidation

Below is the manual consolidation of categories I created.  Primary and secondary categories are the same.

```{r}
#| label: tbl-reasons-mapping
#| tbl-cap: "Reasons"

df_reasons_primary |>
  gt() |>
  tab_header(md("**reason_primary is a manual consolidation of descr_primary**")) |>
  tab_source_note(md("*Daniel Moul. Source: US Peace Corps via Data Liberation Project*"))

```

<br>

### Resignations in all posts

```{r fig.height=10, fig.width=8}
#| label: fig-all-posts-count
#| fig-cap: "Count of non-medevac resignations (all posts)"
#| fig-height: 10
#| fig-width: 8
#| column: page-right

dta_for_table <- d_totals_post |>
  arrange(desc(count)) |>
  mutate(idx = row_number(),
         layout_col = floor(idx / 26) + 1,
         layout_label = glue("Part {layout_col}"),
         post = fct_reorder(post, count)
         )

dta_for_table |>
  select(-layout_col) |>
  ggplot() +
  geom_col(aes(count, post, fill = region),
           alpha = 0.7) +
  scale_fill_viridis_d(end = 0.9) +
  facet_wrap(~layout_label, scales = "free_y") +
  theme(legend.position = "bottom") +
  labs(
    title = "Count of Peace Corps resignations",
    subtitle = "All posts ordered by count; through 2023-07-31; start of date interval is unknown",
    x = NULL,
    y = NULL,
    caption = my_caption
  )

```

<br>

```{r fig.height=10, fig.width=8}
#| label: fig-all-posts-count-sorted-byregion
#| fig-cap: "Count of resignations (all posts by region)"
#| fig-height: 10
#| fig-width: 8
#| column: page-right

dta_for_table <- d_totals_post |>
  mutate(post = fct_reorder(post, count))

dta_for_table |>
  ggplot() +
  geom_col(aes(count, post, fill = region),
           alpha = 0.7) +
  scale_fill_viridis_d(end = 0.9) +
  facet_wrap(~region, scales = "free_y") +
  theme(legend.position = "bottom") +
  labs(
    title = "Count of Peace Corps resignations",
    subtitle = "All posts ordered by count by region; through 2023-07-31; start of date interval is unknown",
    x = NULL,
    y = NULL,
    caption = my_caption
  )

```

<br>

### Other notes

I excluded less than 20 records related to the post "pcr us relief", presumably [Peace Corps Response](https://www.peacecorps.gov/volunteer/peace-corps-response/), for since the data is either not associated with one geographical location or "us relief" implies these assignments were in the USA.

<br>
